Modern BTree\\ techniques

### 2020-01-18 btree/presentation.tex
bw tree

\tikzset{rum-node/.style={draw,circle,fill=white,minimum width=2cm}}
\tikzset{rum-extra-node/.style={rectangle, draw, dashed, rounded corners, fill=white }}
\tikzset{btree-key/.style={minimum height=1cm, pattern=north west lines, draw}}
\tikzset{btree-pointer/.style={minimum height=1cm, draw}}
\tikzset{btree-hide/.style={draw opacity=0, line width=0, pattern=none}}
\tikzset{btree-empty/.style={btree-pointer, dashed}}
\tikzset{btree-leaf/.style={fill=greenGood}}
\tikzset{btree-branch/.style={fill=gray!20}}
\tikzset{btree-line/.style={line width=0.5pt}}
\tikzset{btree-path/.style={btree-pointer, minimum width=0.8cm, fill=red, opacity=0.5}}
\tikzset{>=latex}
\tikzset{btree-high-key/.style={btree-key, pattern=north west lines, pattern color=red, draw=red}}
\tikzset{btree-key-prefix/.style={btree-key, minimum height=0.2cm, fill=red, opacity=0.5}}
\tikzset{btree-key-compare/.style={btree-key, minimum height=0.6cm, fill=blue, opacity=0.5}}
\tikzset{arrow-pointer/.style={
        single arrow,
        minimum height=1.5cm,
        inner sep=3pt,
        line width=1pt,
        draw,
        color=gray,
        single arrow tip angle=45,
        single arrow head extend=0.1cm
    }
}
\tikzset{btree-indirection/.style={btree-key, minimum width=0.2cm, inner sep=0, pattern=none}}
\tikzset{btree-page-header/.style={btree-key, minimum width=0.5cm, inner sep=0, pattern=none}}
\tikzset{cpu-cache/.style={draw, minimum width=1cm, minimum height=1cm, text width=1cm, align=center, rounded corners}}
\tikzset{cpu-cache-hide/.style={cpu-cache, draw=none, color=white}}
\tikzset{btree-node/.style={draw, inner sep=0, minimum width=1cm, minimum height=1cm, rounded corners}}
\tikzset{btree-partition1/.style={draw, inner sep=0, minimum width=1cm, minimum height=1cm, rounded corners, fill=red!20}}
\tikzset{btree-partition2/.style={draw, inner sep=0, minimum width=1cm, minimum height=1cm, rounded corners, fill=green!20}}
\tikzset{btree-partition3/.style={draw, inner sep=0, minimum width=1cm, minimum height=1cm, rounded corners, fill=blue!20}}
\tikzset{smallbtree-node/.style={btree-node, minimum width=0.4cm, minimum height=0.4cm, rounded corners=0.1cm}}
\tikzset{box/.style={draw, dashed, rounded corners, behind path}}
\tikzset{green-box/.style={box, fill=greenGood}}
\tikzset{blue-box/.style={box, fill=blue!20}}
\tikzset{bloom-filter/.style={
    draw,
    isosceles triangle,
    minimum width=1.2cm,
    shape border rotate=-90,
    isosceles triangle apex angle=60,
    rounded corners
}}

@@

@@

    \begin{overprint}[12cm]
        \onslide<1>
        \begin{tikzpicture}[]
            \node[btree-node] (node1) {};

@@

            \draw[->, btree-line, color=white] (node1) -- (node2);
            \draw[->, btree-line, color=white] (node1) -- (node3);
            \draw[->, btree-line, color=white] (node1) -- (node4);

@@

            \draw[->, btree-line, color=white] (node3) -- (node7);
            \draw[->, btree-line, color=white] (node3) -- (node8);

@@

            \draw[->, btree-line] (row21.west)
            .. controls ([xshift=-2cm] row21) and ([xshift=2cm] node8) ..
            (node8.center);

        \end{tikzpicture}
    \end{overprint}
    \end{center}
\end{frame}

>>>>

\tikzset{rum-node/.style={draw,circle,fill=white,minimum width=2cm}}
\tikzset{rum-extra-node/.style={rectangle, draw, dashed, rounded corners, fill=white }}
\tikzset{btree-key/.style={minimum height=1cm, pattern=north west lines, draw}}
\tikzset{btree-pointer/.style={minimum height=1cm, draw}}
\tikzset{btree-hide/.style={draw opacity=0, line width=0, pattern=none}}
\tikzset{btree-empty/.style={btree-pointer, dashed}}
\tikzset{btree-leaf/.style={fill=greenGood}}
\tikzset{btree-branch/.style={fill=gray!20}}
\tikzset{btree-line/.style={line width=0.5pt}}
\tikzset{btree-path/.style={btree-pointer, minimum width=0.8cm, fill=red, opacity=0.5}}
\tikzset{>=latex}
\tikzset{btree-high-key/.style={btree-key, pattern=north west lines, pattern color=red, draw=red}}
\tikzset{btree-key-prefix/.style={btree-key, minimum height=0.2cm, fill=red, opacity=0.5}}
\tikzset{btree-key-compare/.style={btree-key, minimum height=0.6cm, fill=blue, opacity=0.5}}
\tikzset{arrow-pointer/.style={
        single arrow,
        minimum height=1.5cm,
        inner sep=3pt,
        line width=1pt,
        draw,
        color=gray,
        single arrow tip angle=45,
        single arrow head extend=0.1cm
    }
}
\tikzset{btree-indirection/.style={btree-key, minimum width=0.2cm, inner sep=0, pattern=none}}
\tikzset{btree-page-header/.style={btree-key, minimum width=0.5cm, inner sep=0, pattern=none}}
\tikzset{cpu-cache/.style={draw, minimum width=1cm, minimum height=1cm, text width=1cm, align=center, rounded corners}}
\tikzset{cpu-cache-hide/.style={cpu-cache, draw=none, color=white}}
\tikzset{btree-node/.style={draw, inner sep=0, minimum width=1cm, minimum height=1cm, rounded corners}}
\tikzset{btree-leaf-node/.style={btree-node, fill=greenGood}}
\tikzset{btree-branch-node/.style={btree-node, fill=gray!20}}
\tikzset{btree-partition1/.style={draw, inner sep=0, minimum width=1cm, minimum height=1cm, rounded corners, fill=red!20}}
\tikzset{btree-partition2/.style={draw, inner sep=0, minimum width=1cm, minimum height=1cm, rounded corners, fill=green!20}}
\tikzset{btree-partition3/.style={draw, inner sep=0, minimum width=1cm, minimum height=1cm, rounded corners, fill=blue!20}}
\tikzset{smallbtree-node/.style={btree-node, minimum width=0.4cm, minimum height=0.4cm, rounded corners=0.1cm}}
\tikzset{box/.style={draw, dashed, rounded corners, behind path}}
\tikzset{green-box/.style={box, fill=greenGood}}
\tikzset{blue-box/.style={box, fill=blue!20}}
\tikzset{bloom-filter/.style={
    draw,
    isosceles triangle,
    minimum width=1.2cm,
    shape border rotate=-90,
    isosceles triangle apex angle=60,
    rounded corners
}}
\tikzset{bw-delta/.style={draw, fill=redBad, rounded corners}}
\tikzset{bw-page/.style={draw, fill=greenGood, rounded corners}}

@@

    \end{center}
\end{frame}
\note{
    Another interesting approach is partial value on a page and the rest is on
    extended pages (MySQL, SQLite).
}

@@

    \begin{overprint}[12cm]
        \onslide<1>
        \begin{tikzpicture}[]
            \node[btree-branch-node] (node1) {};

            \node[btree-branch-node, below=0.5cm of node1.south west, xshift=-2.5cm] (node2) {};
            \node[btree-branch-node, below=0.5cm of node1.south] (node3) {};
            \node[btree-branch-node, below=0.5cm of node1.south east, xshift=2.5cm] (node4) {};

            \node[btree-leaf-node, below=0.5cm of node2.south west, xshift=-0.2cm] (node5) {};
            \node[btree-leaf-node, below=0.5cm of node2.south east, xshift=0.2cm] (node6) {};

            \node[btree-leaf-node, below=0.5cm of node3.south west, xshift=-0.2cm] (node7) {};
            \node[btree-leaf-node, below=0.5cm of node3.south east, xshift=0.2cm] (node8) {};

            \node[btree-leaf-node, below=0.5cm of node4.south west, xshift=-0.2cm] (node9) {};
            \node[btree-leaf-node, below=0.5cm of node4.south east, xshift=0.2cm] (node10) {};

@@

        \onslide<2>
        \begin{tikzpicture}[]
            \node[btree-branch-node] (node1) {};
            \coordinate[right=0.25cm of node1.center] (link-anchor1);
            \node[below=-0.05cm of link-anchor1] (link1) {\footnotesize{1}};

            \node[btree-branch-node, below=0.5cm of node1.south west, xshift=-2.5cm] (node2) {};
            \node[btree-branch-node, below=0.5cm of node1.south] (node3) {};
            \coordinate[right=0.25cm of node3.center] (link-anchor2);
            \node[below=-0.05cm of link-anchor2] (link2) {\footnotesize{2}};

            \node[btree-branch-node, below=0.5cm of node1.south east, xshift=2.5cm] (node4) {};

            \node[btree-leaf-node, below=0.5cm of node2.south west, xshift=-0.2cm] (node5) {};
            \node[btree-leaf-node, below=0.5cm of node2.south east, xshift=0.2cm] (node6) {};

            \node[btree-leaf-node, below=0.5cm of node3.south west, xshift=-0.2cm] (node7) {};
            \node[btree-leaf-node, below=0.5cm of node3.south east, xshift=0.2cm] (node8) {};

            \node[btree-leaf-node, below=0.5cm of node4.south west, xshift=-0.2cm] (node9) {};
            \node[btree-leaf-node, below=0.5cm of node4.south east, xshift=0.2cm] (node10) {};

            \draw[->, btree-line, color=white] (node1) -- (node2);
            \draw[->, btree-line, color=white] (node1) -- (node3);
            \draw[->, btree-line, dashed] (node1) -- (node4);

@@

            \draw[->, btree-line, color=white] (node3) -- (node7);
            \draw[->, btree-line, dashed] (node3) -- (node8);

@@

            \draw[->, btree-line] (row21.west)
            .. controls ([xshift=-2cm] row21) and ([xshift=2cm] node8) ..
            (node8.center);
        \end{tikzpicture}

        \onslide<3>
        \begin{tikzpicture}[]
            \path
                node[draw, right=5cm of node1.east, minimum width=1.2cm] (header1) {\textbf{\normalsize{ID}}}
                node[draw, right=0 of header1, minimum width=1.2cm] (header2) {\textbf{\normalsize{PTR}}}
                    node[draw, below=0 of header1, minimum width=1.2cm] (row11) {\normalsize{1}}
                    node[draw, right=0 of row11, minimum width=1.2cm] (row12) {\normalsize{0x...}}
                    node[draw, below=0 of row11, minimum width=1.2cm] (row21) {\normalsize{2}}
                    node[draw, right=0 of row21, minimum width=1.2cm] (row22) {\normalsize{0x...}}
                    node[behind path, fit=(header1)(header2)(row11)(row12)(row21)(row22)] (mapping) {};

            \node[bw-delta, right=4cm of header2, color=white] (delta) {$\Delta$};
            \node[bw-page, below=0.5cm of delta] (page) {Page};
            \draw[->, btree-line] (row22)
                .. controls ([yshift=-2cm] row22) and ([yshift=-2cm] page) ..
            (page);
        \end{tikzpicture}

        \onslide<4>
        \begin{tikzpicture}[]
            \path
                node[draw, right=5cm of node1.east, minimum width=1.2cm] (header1) {\textbf{\normalsize{ID}}}
                node[draw, right=0 of header1, minimum width=1.2cm] (header2) {\textbf{\normalsize{PTR}}}
                    node[draw, below=0 of header1, minimum width=1.2cm] (row11) {\normalsize{1}}
                    node[draw, right=0 of row11, minimum width=1.2cm] (row12) {\normalsize{0x...}}
                    node[draw, below=0 of row11, minimum width=1.2cm] (row21) {\normalsize{2}}
                    node[draw, right=0 of row21, minimum width=1.2cm] (row22) {\normalsize{0x...}}
                    node[behind path, fit=(header1)(header2)(row11)(row12)(row21)(row22)] (mapping) {};

            \node[bw-delta, right=4cm of header2] (delta) {$\Delta$};
            \node[bw-page, below=0.5cm of delta] (page) {Page};
            \draw[->, btree-line] (delta) -- (page);
            \draw[->, btree-line] (row22)
                .. controls ([yshift=-2cm] row22) and ([yshift=-2cm] page) ..
            (page);
        \end{tikzpicture}

        \onslide<5>
        \begin{tikzpicture}[]
            \path
                node[draw, right=5cm of node1.east, minimum width=1.2cm] (header1) {\textbf{\normalsize{ID}}}
                node[draw, right=0 of header1, minimum width=1.2cm] (header2) {\textbf{\normalsize{PTR}}}
                    node[draw, below=0 of header1, minimum width=1.2cm] (row11) {\normalsize{1}}
                    node[draw, right=0 of row11, minimum width=1.2cm] (row12) {\normalsize{0x...}}
                    node[draw, below=0 of row11, minimum width=1.2cm] (row21) {\normalsize{2}}
                    node[draw, right=0 of row21, minimum width=1.2cm] (row22) {\normalsize{0x...}}
                    node[behind path, fit=(header1)(header2)(row11)(row12)(row21)(row22)] (mapping) {};

            \node[bw-delta, right=4cm of header2] (delta) {$\Delta$};
            \node[bw-page, below=0.5cm of delta] (page) {Page};
            \draw[->, btree-line] (delta) -- (page);
            \draw[->, btree-line, dashed, color=gray] (row22)
                .. controls ([yshift=-2cm] row22) and ([yshift=-2cm] page) ..
                node[midway, above, color=gray] {\footnotesize{CAS}} (page);
            \draw[->, btree-line] (row22)
                .. controls ([xshift=2cm] row22) and ([xshift=-2cm] delta) ..
            (delta);
        \end{tikzpicture}

        \onslide<6>
        \begin{tikzpicture}[]
            \path
                node[draw, right=5cm of node1.east, minimum width=1.2cm] (header1) {\textbf{\normalsize{ID}}}
                node[draw, right=0 of header1, minimum width=1.2cm] (header2) {\textbf{\normalsize{PTR}}}
                    node[draw, below=0 of header1, minimum width=1.2cm] (row11) {\normalsize{1}}
                    node[draw, right=0 of row11, minimum width=1.2cm] (row12) {\normalsize{0x...}}
                    node[draw, below=0 of row11, minimum width=1.2cm] (row21) {\normalsize{2}}
                    node[draw, right=0 of row21, minimum width=1.2cm] (row22) {\normalsize{0x...}}
                    node[behind path, fit=(header1)(header2)(row11)(row12)(row21)(row22)] (mapping) {};

            \node[bw-delta, right=5cm of header2] (delta1) {$\Delta_1$};
            \node[bw-delta, left=0.5cm of delta1] (delta2) {$\Delta_2$};
            \node[bw-delta, left=0.5cm of delta2] (delta3) {$\Delta_3$};
            \node[bw-page, below=0.5cm of delta2] (page) {Page};
            \node[behind path, fit=(delta1)(delta2)(delta3)(page), rounded corners, color=white] (merged) {};
            \draw[->, btree-line] (delta1) -- (delta2) -- (delta3) -- (page);
            \draw[->, btree-line] (row22)
                .. controls ([xshift=2cm] row22) and ([xshift=-2cm] delta3) ..
            (delta3);
        \end{tikzpicture}

        \onslide<7>
        \begin{tikzpicture}[]
            \path
                node[draw, right=5cm of node1.east, minimum width=1.2cm] (header1) {\textbf{\normalsize{ID}}}
                node[draw, right=0 of header1, minimum width=1.2cm] (header2) {\textbf{\normalsize{PTR}}}
                    node[draw, below=0 of header1, minimum width=1.2cm] (row11) {\normalsize{1}}
                    node[draw, right=0 of row11, minimum width=1.2cm] (row12) {\normalsize{0x...}}
                    node[draw, below=0 of row11, minimum width=1.2cm] (row21) {\normalsize{2}}
                    node[draw, right=0 of row21, minimum width=1.2cm] (row22) {\normalsize{0x...}}
                    node[behind path, fit=(header1)(header2)(row11)(row12)(row21)(row22)] (mapping) {};

            \path
                node[bw-delta, right=5cm of header2] (delta1) {$\Delta_1$}
                node[bw-delta, left=0.5cm of delta1] (delta2) {$\Delta_2$}
                node[bw-delta, left=0.5cm of delta2] (delta3) {$\Delta_3$}
                node[bw-page, below=0.5cm of delta2] (page) {Page}
                node[behind path, fill=gray!20, fit=(delta1)(delta2)(delta3)(page), rounded corners] (merged) {};
            \draw[->, btree-line] (delta1) -- (delta2) -- (delta3) -- (page);
            \draw[->, btree-line] (row22)
                .. controls ([xshift=2cm] row22) and ([xshift=-2cm] delta3) ..
            (delta3);

            \node[bw-page, below=1cm of page] (merged-page) {Merged Page};
            \draw[->, btree-line] (merged.south) -- (merged-page);
        \end{tikzpicture}

        \onslide<8>
        \begin{tikzpicture}[]
            \path
                node[draw, right=5cm of node1.east, minimum width=1.2cm] (header1) {\textbf{\normalsize{ID}}}
                node[draw, right=0 of header1, minimum width=1.2cm] (header2) {\textbf{\normalsize{PTR}}}
                    node[draw, below=0 of header1, minimum width=1.2cm] (row11) {\normalsize{1}}
                    node[draw, right=0 of row11, minimum width=1.2cm] (row12) {\normalsize{0x...}}
                    node[draw, below=0 of row11, minimum width=1.2cm] (row21) {\normalsize{2}}
                    node[draw, right=0 of row21, minimum width=1.2cm] (row22) {\normalsize{0x...}}
                    node[behind path, fit=(header1)(header2)(row11)(row12)(row21)(row22)] (mapping) {};

            \path
                node[bw-delta, right=5cm of header2] (delta1) {$\Delta_1$}
                node[bw-delta, left=0.5cm of delta1] (delta2) {$\Delta_2$}
                node[bw-delta, left=0.5cm of delta2] (delta3) {$\Delta_3$}
                node[bw-page, below=0.5cm of delta2] (page) {Page}
                node[behind path, fill=gray!20, fit=(delta1)(delta2)(delta3)(page), rounded corners] (merged) {};
            \draw[->, btree-line] (delta1) -- (delta2) -- (delta3) -- (page);
            \draw[->, btree-line, color=gray, dashed] (row22)
                .. controls ([xshift=2cm] row22) and ([xshift=-2cm] delta3) ..
            node[midway, above, color=gray] {\footnotesize{CAS}} (delta3);

            \node[bw-page, below=1cm of page, rounded corners] (merged-page) {Merged Page};
            \draw[->, btree-line] (merged.south) -- (merged-page);
            \draw[->, btree-line] (row22)
                .. controls ([xshift=2cm] row22) and ([xshift=-3cm] merged-page) ..
            (merged-page.west);
        \end{tikzpicture}

    \end{overprint}
    \end{center}
\end{frame}
\note{
    Requires also "split", "index", "remove node", "merge", "flush" deltas
}
